\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-06-18 doc/USER_MANUAL/02_getting_started.tex
updates user manual

The SCOTCH library \citep{PeRo96} provides efficent static mapping,
graph and mesh partitioning routines. SCOTCH is a free software package
developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
In case no SCOTCH libraries can be found on the system, the configuration
will bundle the version provided with the source code for compilation.
The path to an existing SCOTCH installation can to be set explicitly
with the option \texttt{-{}-with-scotch-dir}. Just as an example:
\begin{verbatim}
./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch
\end{verbatim}
\noindent
If you use the Intel ifort compiler to compile the code, we recommend
that you use the Intel icc C compiler to compile Scotch, i.e., use:
\begin{verbatim}
./configure CC=icc FC=ifort MPIFC=mpif90
\end{verbatim}

@@

A summary of the most important configuration variables follows.
\begin{description}
\item [{\texttt{F90}}] Path to the Fortran compiler.
\item [{\texttt{MPIF90}}] Path to MPI Fortran.
\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
MPI libraries.
\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
\end{description}
The configuration script automatically creates for each executable
a corresponding \texttt{Makefile} in the \texttt{src/} subdirectoy.
The \texttt{Makefile} contains a number of suggested entries for various
compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
has run on a wide variety of compute platforms, e.g., various PC clusters
and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
you wish to use on your system and choose the related optimization
flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
not optimal for your system, so we encourage you to experiment with
these flags and to solicit advice from your systems administrator.
Selecting the right compiler and optimization flags can make a tremendous
difference in terms of performance. We welcome feedback on your experience
with various compilers and flags.

@@

Regular POSIX I/O can be problematic when dealing with large simulations one large
clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
to deal transparently take advantage of advanced paralllel file system features. To enable
ADIOS, the following steps should be done:
\begin{enumerate}
\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
Make sure that your environment variables reference it.
\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
The default values probably suit most cases.
\item Configure using the \texttt{--with-adios} flag.
\end{enumerate}
ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.

>>>>

The SCOTCH library \citep{PeRo96} provides efficient static mapping,
graph and mesh partitioning routines. SCOTCH is a free software package
developed by Fran\c{c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux,
France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
In case no SCOTCH libraries can be found on the system, the configuration
will bundle the version provided with the source code for compilation.
The path to an existing SCOTCH installation can to be set explicitly
with the option \texttt{-{}-with-scotch-dir}. Just as an example:
\begin{verbatim}
./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch
\end{verbatim}
\noindent
If you use the Intel ifort compiler to compile the code, we recommend
that you use the Intel icc C compiler to compile Scotch, i.e., use:
\begin{verbatim}
./configure CC=icc FC=ifort MPIFC=mpif90
\end{verbatim}

@@

A summary of the most important configuration variables follows.
\begin{description}
\item [{\texttt{F90}}] Path to the Fortran compiler.
\item [{\texttt{MPIF90}}] Path to MPI Fortran.
\item [{\texttt{MPI\_FLAGS}}] Some systems require this flag to link to
MPI libraries.
\item [{\texttt{FLAGS\_CHECK}}] Compiler flags.
\end{description}
The configuration script automatically creates for each executable
a corresponding \texttt{Makefile} in the \texttt{src/} subdirectory.
The \texttt{Makefile} contains a number of suggested entries for various
compilers, e.g., Portland, Intel, Absoft, NAG, and Lahey. The software
has run on a wide variety of compute platforms, e.g., various PC clusters
and machines from Sun, SGI, IBM, Compaq, and NEC. Select the compiler
you wish to use on your system and choose the related optimization
flags. Note that the default flags in the \texttt{Makefile} are undoubtedly
not optimal for your system, so we encourage you to experiment with
these flags and to solicit advice from your systems administrator.
Selecting the right compiler and optimization flags can make a tremendous
difference in terms of performance. We welcome feedback on your experience
with various compilers and flags.

@@

Regular POSIX I/O can be problematic when dealing with large simulations one large
clusters (typically more than $10,000$ processes). SPECFEM3D use the ADIOS library~\cite{Liu2013}
to deal transparently take advantage of advanced parallel file system features. To enable
ADIOS, the following steps should be done:
\begin{enumerate}
\item Install ADIOS (available from \url{https://www.olcf.ornl.gov/center-projects/adios/}).
Make sure that your environment variables reference it.
\item You may want to change ADIOS related values in the \texttt{constants.h.in} file.
The default values probably suit most cases.
\item Configure using the \texttt{--with-adios} flag.
\end{enumerate}
ADIOS is currently only usable for meshfem3D generated mesh (i.e. not for meshes generated
with CUBTI). Additional control parameters are discussed in section~\ref{cha:Main-Parameter}.

### 2015-06-18 doc/USER_MANUAL/04_creating_databases.tex
updates user manual

\item [{\texttt{external}}] For a user-specified isotropic model which
uses externally defined model parameters. Uses external model definitions
set up in routines located in file \texttt{model\_external\_values.f90}
in directory \texttt{src/generate\_databases/}. Please modify these
generic template routines to use your own model definitions.
\item [{\texttt{gll}}] For a user-specified isotropic model which uses
external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
are given in the same format as when outputted by the \texttt{xgenerate\_databases}
executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
files define the model parameters on all GLL points which can be used
for iterative inversion procedures.
\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
in directory \texttt{DATA/st\_3D\_block\_harvard/}.
\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
a tomographic model file \texttt{tomographic\_model.xyz} in directory
\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
on how to specify your own 3D tomographic model.

@@

TODO / future work to do: currently the \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL} 
option assumes to have the (master) mesh files in run0001/DATABASES\_MPI or run0001/OUTPUT\_FILES/DATABASES\_MPI. 
However, for adjoint runs you still need a DATABASES\_MPI folder in each of the sub-runs directories,
e.g. run0002/DATABASES\_MPI, etc. to store the forward wavefields, kernels etc. of each sub-run.
This would not be needed for forward simulations.

@@

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name nummaterial\_poroelastic\_file, which is located in the directory MESH, the format is as follow:
\begin{verbatim}
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}
where\\
\texttt{rho\_s} = solid density,\\
\texttt{rho\_f} = fluid density,\\
\texttt{phi} = porosity,\\
\texttt{tort} = tortuosity,\\
\texttt{kxx} = xx component of permeability tensor,\\
\texttt{kxy} = xy,yx components of permeability tensor,\\
\texttt{kyy} = yy component of permeability tensor,\\
\texttt{kxz} = xz,zx components of permeability tensor,\\
\texttt{kzz} = zz component of permeability tensor,\\
\texttt{kappa\_s} = solid bulk modulus,\\
\texttt{kappa\_f} = fluid bulk modulus,\\
\texttt{kappa\_fr} = frame bulk modulus,\\
\texttt{eta\_f} = fluid viscosity,\\
\texttt{mu\_fr} = frame shear modulus.\\

Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is elastic, 2 is acoustic ? see the example in the package: examples/homogeneous\_poroelastic.

>>>>

\item [{\texttt{external}}] For a user-specified isotropic model which
uses externally defined model parameters. Uses external model definitions
set up in routines located in file \texttt{model\_external\_values.f90}
in directory \texttt{src/generate\_databases/}. Please modify these
generic template routines to use your own model definitions.

\item [{\texttt{gll}}] For a user-specified isotropic model which uses
external binary files for $v_{p}$, $v_{s}$ and $\rho$. Binary files
are given in the same format as when outputted by the \texttt{xgenerate\_databases}
executable when using option \texttt{SAVE\_MESH\_FILES}. These binary
files define the model parameters on all GLL points which can be used
for iterative inversion procedures. Note that for simulation setups with attenuation, 
it will also read in the external binary mesh files for $Q_{\kappa}$ and $Q_{\mu}$.

\item [{\texttt{salton\_trough}}] A 3D $V_{p}$ model for Southern California.
Users must provide the corresponding data file \texttt{regrid3\_vel\_p.bin}
in directory \texttt{DATA/st\_3D\_block\_harvard/}.

\item [{\texttt{tomo}}] For a user-specified 3D isotropic model which uses
a tomographic model file \texttt{tomographic\_model.xyz} in directory
\texttt{DATA}. See Section \ref{sec:Using-tomographic}, for a discussion
on how to specify your own 3D tomographic model.
\end{description}
\end{description}

@@

TODO / future work to do: currently the \texttt{BROADCAST\_SAME\_MESH\_AND\_MODEL}
option assumes to have the (master) mesh files in \texttt{run0001/DATABASES\_MPI} or 
\texttt{run0001/OUTPUT\_FILES/DATABASES\_MPI}.
However, for adjoint runs you still need a \texttt{DATABASES\_MPI/} folder in each of the sub-runs directories,
e.g. \texttt{run0002/DATABASES\_MPI}, etc. to store the forward wavefields, kernels etc. of each sub-run.
This would not be needed for forward simulations.

@@

The way we prescribe material property for porous material in SPECFEM3D is as follow:
We use a file name "nummaterial\_poroelastic\_file", which is located in the directory \texttt{MESH/}, the format is as follow:
\begin{verbatim}
 rhos rhof phi c kxx kxy kxz kyy kyz kzz Ks Kf Kfr etaf mufr
\end{verbatim}
where\\
\texttt{rho\_s} = solid density,\\
\texttt{rho\_f} = fluid density,\\
\texttt{phi} = porosity,\\
\texttt{tort} = tortuosity,\\
\texttt{kxx} = xx component of permeability tensor,\\
\texttt{kxy} = xy,yx components of permeability tensor,\\
\texttt{kyy} = yy component of permeability tensor,\\
\texttt{kxz} = xz,zx components of permeability tensor,\\
\texttt{kzz} = zz component of permeability tensor,\\
\texttt{kappa\_s} = solid bulk modulus,\\
\texttt{kappa\_f} = fluid bulk modulus,\\
\texttt{kappa\_fr} = frame bulk modulus,\\
\texttt{eta\_f} = fluid viscosity,\\
\texttt{mu\_fr} = frame shear modulus.\\

Using an external mesh (for instance coming from CUBIT/TRELIS), poroelastic materials have the ID number 3, while 1 is acoustic and 2 is elastic (see the example in the package: EXAMPLES/homogeneous\_poroelastic).

### 2015-06-18 doc/USER_MANUAL/11_graphics.tex
updates user manual

To visualize these files, we use an auxilliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile
it in the root directory, type:

@@

The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
Therefore, we first need to collect these files on the front end,
combine them into one mesh file, and visualize them with some auxilliary
programs.
\begin{enumerate}
\item \textbf{Create slice files}

@@

We use an auxilliary program \texttt{combine\_vol\_data.f90} to combine
the kernel files from all slices into one mesh file.
\begin{enumerate}
\item Compile it in the root directory:
{\footnotesize
\begin{verbatim}
make xcombine_vol_data
./bin/xcombine_vol_data slice_list filename input_dir output_dir high/low-resolution
\end{verbatim}
}
where \texttt{input\_dir} is the directory where all the individual
kernel files are stored, and \texttt{output\_dir} is where the mesh
file will be written.

>>>>

To visualize these files, we use an auxiliary program \texttt{combine\_vol\_data.f90}
to combine the data from all slices into one mesh file. To compile
it in the root directory, type:

@@

The finite-frequency kernels computed as explained in Section \ref{sec:Adjoint-simulation-finite}
are saved in the \texttt{LOCAL\_PATH} at the end of the simulation.
Therefore, we first need to collect these files on the front end,
combine them into one mesh file, and visualize them with some auxiliary
programs.
\begin{enumerate}
\item \textbf{Create slice files}

@@

We use an auxiliary program \texttt{combine\_vol\_data.f90} to combine
the kernel files from all slices into one mesh file.
\begin{enumerate}
\item Compile it in the root directory:
{\footnotesize
\begin{verbatim}
make xcombine_vol_data
./bin/xcombine_vol_data slice_list filename input_dir output_dir high/low-resolution
\end{verbatim}
}
where \texttt{input\_dir} is the directory where all the individual
kernel files are stored, and \texttt{output\_dir} is where the mesh
file will be written.

### 2015-06-18 doc/USER_MANUAL/13_changing_the_model.tex
updates user manual

The external tomographic model is represented by a grid of points
with assigned material properties and homogeneous resolution along
each spatial direction x, y and z. The ASCII file \texttt{file\_name}
that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
directory, set in the \texttt{Par\_file}. The format of the file,
as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
directory, looks like Figure \ref{fig:tomography_file}, where
\begin{description}
\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
of the initial and final tomographic grid points along the x direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
of the initial and final tomographic grid points along the y direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
of the initial and final tomographic grid points along the z direction
(in the mesh units, e.g., $m$);
\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
spacing between the tomographic grid points along the x, y and z directions,
respectively (in the mesh units, e.g., $m$);
\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
along the spatial directions x, y and z, respectively; \texttt{NX}
is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
y and z directions, respectively;
\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
these values could be the actual limits of the tomographic parameters
in the grid or the minimum and maximum values to which we force the
cut of velocity and density in the model.
\end{description}
After the first four lines, in the file \texttt{file\_name} the tomographic
grid points are listed with the corresponding values of \texttt{vp},
\texttt{vs} and \texttt{rho}, scanning the grid along the x coordinate
(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
\texttt{END\_Z}, with step of \texttt{SPACING\_Z}). %fig
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{figures/tomo_file.jpg}
\par\end{centering}

\caption{Tomography file \texttt{file\_name} that describes an external Earth
model. The coordinates x, y and z of the grid, their limits ORIG\_X,
ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
(e.g., UTM coordinates in meters).}

@@

The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions. (Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
%
\begin{verbatim}
2  -1 tomography elastic tomography_model.xyz 1
\end{verbatim}
%
and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}. All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated accordind to the values in \texttt{tomography\_model.xyz}.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
                           rho, vp, vs, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qmu\_atten}}] Shear wave quality factor: $0<Q_{\mu}<9000$
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

@@

To use your anisotropic models, you can either set in the \texttt{Par\_file}
the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
models may be superimposed on the mesh based upon the subroutine in
file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
The call to this subroutine is of the form
\begin{verbatim}
call model_aniso(iflag_aniso, rho, vp, vs, &
                 c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26, &
                 c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
generic pre-defined anisotropic models.
\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
$\rho$, $v_{p}$ and $v_{s}$.
\end{description}
Output from the routine consists of the following non-dimensional
model parameters:
\begin{description}
\item [{\texttt{c11},}] \textbf{$\cdots$,} \texttt{\textbf{c66}} 21~dimensionalized
anisotropic elastic parameters.
\end{description}
You can replace the \texttt{model\_aniso.f90} file by your own version
\textit{provided you do not change the call structure of the routine},
i.e., the new routine should take exactly the same input and produce
the required relative output.

>>>>

The external tomographic model is represented by a grid of points
with assigned material properties and homogeneous resolution along
each spatial direction x, y and z. The ASCII file \texttt{file\_name}
that describe the tomography should be located in the \texttt{TOMOGRAPHY\_PATH}
directory, set in the \texttt{Par\_file}. The format of the file,
as read from \texttt{model\_tomography.f90} located in the \texttt{src/generate\_databases}
directory, looks like Figure \ref{fig:tomography_file}, starting with a header information where
\begin{description}
\item [{\texttt{ORIG\_X}, \texttt{END\_X}}] are, respectively, the coordinates
of the initial and final tomographic grid points along the x direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Y}, \texttt{END\_Y}}] respectively the coordinates
of the initial and final tomographic grid points along the y direction
(in the mesh units, e.g., $m$);
\item [{\texttt{ORIG\_Z}, \texttt{END\_Z}}] respectively the coordinates
of the initial and final tomographic grid points along the z direction
(in the mesh units, e.g., $m$);
\item [{\texttt{SPACING\_X}, \texttt{SPACING\_Y}, \texttt{SPACING\_Z}}] the
spacing between the tomographic grid points along the x, y and z directions,
respectively (in the mesh units, e.g., $m$);
\item [{\texttt{NX}, \texttt{NY}, \texttt{NZ}}] the number of grid points
along the spatial directions x, y and z, respectively; \texttt{NX}
is given by {[}(\texttt{END\_X} - \texttt{ORIG\_X})/\texttt{SPACING\_X}{]}+1;
\texttt{NY} and \texttt{NZ} are the same as \texttt{NX}, but for the
y and z directions, respectively;
\item [{\texttt{VP\_MIN}, \texttt{VP\_MAX}, \texttt{VS\_MIN}, \texttt{VS\_MAX}, \texttt{RHO\_MIN}, \texttt{RHO\_MAX}}] the
minimum and maximum values of the wave speed \texttt{vp} and \texttt{vs}
(in $m\, s^{-1}$) and of the density \texttt{rho} (in $kg\, m^{-3}$);
these values could be the actual limits of the tomographic parameters
in the grid or the minimum and maximum values to which we force the
cut of velocity and density in the model.
\end{description}
After the first four lines, the tomography file \texttt{file\_name} lists the data record where all tomographic
grid points are listed with the corresponding values of \texttt{vp},
\texttt{vs} and \texttt{rho} (and optionally $Q_{p}$ and $Q_{s}$), scanning the grid along the x coordinate
(from \texttt{ORIG\_X} to \texttt{END\_X} with step of \texttt{SPACING\_X})
for each given y (from \texttt{ORIG\_Y} to \texttt{END\_Y}, with step
of \texttt{SPACING\_Y}) and each given z (from \texttt{ORIG\_Z} to
\texttt{END\_Z}, with step of \texttt{SPACING\_Z}).

\noindent
Each data record line provides the velocity model values in a format like:  
{\small
\begin{verbatim}
# data record format: purely isotropic model
x	y	z		vp		vs		rho
..
\end{verbatim}
}
\noindent
or
{\small
\begin{verbatim}
# data record format: anelastic isotropic model
x	y	z		vp		vs		rho	Qp	Qs
..
\end{verbatim}
}
\noindent
where  \texttt{x}, \texttt{y}, \texttt{z} are the grid point position, \texttt{vp}, \texttt{vs} and \texttt{rho} the P- and S-wave speeds and density, and \texttt{Qp} and \texttt{Qs} the quality factors for P- and S-wave speeds. The quality factors are optional and can be omitted for purely elastic models (the tomography routine will recognize both formats). Internally, the quality factors will be converted to bulk and shear attenuation values, 
$Q_{\kappa}$ and $Q_{\mu}$ respectively \citep{AnHa78}. 
For simulations with attenuation, please note that the Vp- and Vs-velocities of your model are given for a reference
frequency. To change this reference frequency, you change the value
of \texttt{ATTENUATION\_f0\_REFERENCE} in the main constants file
\texttt{constants.h} found in subdirectory \texttt{src/shared/}.

%fig%
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=1\textwidth]{figures/tomo_file.jpg}
\par\end{centering}

\caption{Tomography file \texttt{file\_name} that describes an external, purely elastic Earth
model. The coordinates x, y and z of the grid, their limits ORIG\_X,
ORIG\_Y, ORIG\_Z, END\_X, END\_Y, END\_Z and the grid spacings SPACING\_X,
SPACING\_Y, SPACING\_Z should be in the units of the constructed mesh
(e.g., UTM coordinates in meters).}

@@

The example in Figure~\ref{fig:tomography_file} is for an external tomography file with uniform spacing in the $x$, $y$, and $z$ directions. (Note that $\Delta x$, $\Delta y$, and $\Delta z$ need not be the same as in the example.) In that example, the  \texttt{nummaterial\_velocity\_file} is
%
\begin{verbatim}
2  -1 tomography elastic tomography_model.xyz 1
\end{verbatim}
%
and the file \texttt{tomography\_model.xyz} will need to reside in \texttt{DATA/}. All entries in the second column of \texttt{materials\_file} will be \texttt{-1}, which means that each element in the mesh will be interpolated according to the values in \texttt{tomography\_model.xyz}.

@@

To use your own external model, you can set in the \texttt{Par\_file}
the model parameter \texttt{MODEL = external}. Three-dimensional acoustic
and/or (an)elastic (attenuation) models may then be superimposed onto
the mesh based upon your subroutine in \texttt{model\_external\_values.f90}
located in directory \texttt{src/generate\_databases}. The call to
this routine would be as follows
\begin{verbatim}
call model_external_values(xmesh, ymesh, zmesh, &
            rho, vp, vs, qkappa_atten, qmu_atten, iflag_aniso, idomain_id)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{xmesh,ymesh,zmesh}}] location of mesh point
\end{description}
Output to this routine consists of:
\begin{description}
\item [{\texttt{rho,vp,vs}}] isotropic model parameters for density $\rho$
($kg/m^{3}$), $v_{p}$ ($m/s$) and $v_{s}$ ($m/s$)
\item [{\texttt{qkappa\_atten,qmu\_atten}}] Bulk and shear wave quality factor: $0<Q_{\kappa,\mu}<9000$
\item [{\texttt{iflag\_aniso}}] anisotropic model flag, $0$ indicating
no anisotropy or $1$ using anisotropic model parameters as defined
in routine file \texttt{model\_aniso.f90}
\item [{\texttt{idomain\_id}}] domain identifier, $1$ for acoustic, $2$
for elastic, $3$ for poroelastic materials.
\end{description}
Note that the resolution and maximum value of anelastic models are
truncated. This speeds the construction of the standard linear solids
during the meshing stage. To change the resolution, currently at one
significant figure following the decimal, or the maximum value (9000),
consult \texttt{shared/constants.h}.

@@

To use your anisotropic models, you can either set in the \texttt{Par\_file}
the model parameter \texttt{ANISOTROPY} to \texttt{.true.} or set
the model parameter \texttt{MODEL = aniso}. Three-dimensional anisotropic
models may be superimposed on the mesh based upon the subroutine in
file \texttt{model\_aniso.f90} located in directory \texttt{src/generate\_databases}.
The call to this subroutine is of the form
\begin{verbatim}
call model_aniso(iflag_aniso, rho, vp, vs, &
                 c11,c12,c13,c14,c15,c16,c22,c23,c24,c25,c26, &
                 c33,c34,c35,c36,c44,c45,c46,c55,c56,c66)
\end{verbatim}
Input to this routine consists of:
\begin{description}
\item [{\texttt{iflag\_aniso}}] flag indicating the type of the anisotropic
model, i.e. $0$ for no superimposed anisotropy, $1$ or $2$ for
generic pre-defined anisotropic models.
\item [{\texttt{rho,vp,vs}}] reference isotropic model parameters for density
$\rho$, $v_{p}$ and $v_{s}$.
\end{description}
Output from the routine consists of the following non-dimensional
model parameters:
\begin{description}
\item [{\texttt{c11,..,c66}}] 21~dimensionalized
anisotropic elastic parameters.
\end{description}
You can replace the \texttt{model\_aniso.f90} file by your own version
\textit{provided you do not change the call structure of the routine},
i.e., the new routine should take exactly the same input and produce
the required relative output.
